A Comparision of Bootstrap Confidence Interval Methods

### 2022-12-03 project/PaperTemplate/paper.Rmd
Finish first draft

---
title: A Comparision of Bootstrap Methods

@@

keywords:
- confidence interval
- simulation
- percentile
- bias-corrected
- acceleration
- studentized

@@

```{r}
data(SnowGR)
favstats(~ Dec, data = SnowGR)
gf_histogram(~ Dec, data = SnowGR) +
  labs(x = "Annual Snowfall in Decemeber", 
       title = "Distribution of Annual Snowfall in Decemeber", y = "Count")
```

@@

```{r}
orig_mean <- mean(~ Dec, data = SnowGR) # theta hat

@@

```{r}
gf_dhistogram(~ mean, data = dec_means) %>%
  gf_dens() %>%
  gf_labs(title = "Histogram of 10,000 Bootstrapped Mean Dec Values")
```

@@

```{r}
mean(~ Dec, data = SnowGR) + qnorm(c(0.025, 0.975)) * 
  sd(~ Dec, data = SnowGR)/sqrt(nrow(SnowGR))
```

```{r}
gf_histogram(~ Dec, data = SnowGR) +
  labs(x = "Annual Snowfall in Decemeber", 
       title = "Distribution of Annual Snowfall in Decemeber", y = "Count") +
  geom_vline(aes(xintercept= 13.85639)) +
  geom_vline(aes(xintercept= 17.65957))
```

@@

```{r}
qdata(~ mean, c(0.025, 0.975), data = dec_means)
```

```{r}
gf_histogram(~ Dec, data = SnowGR) +
  labs(x = "Annual Snowfall in Decemeber", 
       title = "Distribution of Annual Snowfall in Decemeber", y = "Count") +
  geom_vline(aes(xintercept= 13.91170)) +
  geom_vline(aes(xintercept= 17.72353))
```
This interval is similar to the one created using the Standard method, but shifted slightly to the right.

@@

```{r}
sample_mean <- mean(~Dec, data = SnowGR); sample_mean
```

Then we find the proportion of bootstrap replications that have a sample mean less than the original sample mean.
```{r}
less_than_sample_mean <- sum(ifelse(dec_means <= sample_mean, 1, 0))/10000
less_than_sample_mean
```

From this, we can calculate the bias-correction value, $z_0$:
```{r}
z0 <- qnorm(less_than_sample_mean); z0
```

@@

```{r}
alphalow <- 0.025; alphahigh <- 0.975

@@

```{r}
gf_histogram(~ Dec, data = SnowGR) +
  labs(x = "Annual Snowfall in Decemeber", 
       title = "Distribution of Annual Snowfall in Decemeber", y = "Count") +
  geom_vline(aes(xintercept= 14.01172)) +
  geom_vline(aes(xintercept= 17.82017))

@@

```{r}
theta <- function(x){mean(x)}
jackknife_results <- bootstrap::jackknife(SnowGR$Dec, theta) 
```

@@

```{r}
jackmean <- mean(~ jackknife_results$jack.values); jackmean
```

@@

```{r}
estimated_a <- (1/6)*sum((jackknife_results$jack.values - jackmean)^3)/
  (sum((jackknife_results$jack.values - jackmean)^2))^(1.5); estimated_a
```

@@

```{r}
a_newlow <- pnorm(z0 + (z0 + qnorm(alphalow))/
                    (1-estimated_a*(z0 + qnorm(alphalow)))); #a_newlow
a_newhigh <- pnorm(z0 + (z0 + qnorm(alphahigh))/
                     (1-estimated_a*(z0 + qnorm(alphahigh)))); #a_newhigh
qdata(~ mean, c(a_newlow, a_newhigh), data = dec_means)
```

```{r}
gf_histogram(~ Dec, data = SnowGR) +
  labs(x = "Annual Snowfall in Decemeber", 
       title = "Distribution of Annual Snowfall in Decemeber", y = "Count") +
  geom_vline(aes(xintercept= 13.91591)) +
  geom_vline(aes(xintercept= 17.72353))

@@

```{r}
set.seed(495)

@@

@@

Third, we sample from a Log-Normal distribution with mean log of 0 and a mean standard deviation of 1. Therefore, the true population mean of this distribution is $e^{\frac{1}{2}}$ [@Guillaume21].

```{r run simulation log normal}
run_simulation_log_normal <- function(simulation_rep){
  trueparameter <- exp(1/2)

@@

Lastly, we sample from a Beta distribution with shape parameters of 4 and 1 which gives us a true population mean of 0.8.

```{r run simulation beta}
run_simulation_beta <- function(simulation_rep){
  trueparameter <- 0.8

@@

```{r collect data gamma}
collect_data_gamma <- function(sim_reps) {
  data2 <- matrix(nrow = 5, ncol = 6)
  sample_sizes = 50
  for (i in 1:5) {
    simulation_rep <- seq(1,sim_reps)
    results <- mclapply(simulation_rep, run_simulation_gamma, mc.cores = detectCores())
    df <- data.frame(matrix(unlist(results), nrow=length(results), byrow=TRUE))
    ci_prop <- unname(colSums(df)/sim_reps)

    data2[i, 1] <- sample_sizes
    data2[i, 2:6] <- ci_prop
    sample_sizes <- sample_sizes*2
  }

@@

```{r collect data log normal}
collect_data_log_normal <- function(sim_reps) {
  data2 <- matrix(nrow = 5, ncol = 6)
  sample_sizes = 50

@@

    data2[i, 1] <- sample_sizes
    data2[i, 2:6] <- ci_prop
    sample_sizes <- sample_sizes*2
  }

@@

```{r collect data normal}
collect_data_normal <- function(sim_reps) {
  data2 <- matrix(nrow = 5, ncol = 6)
  sample_sizes = 50

@@

```{r collect data beta}
collect_data_beta <- function(sim_reps) {
  data2 <- matrix(nrow = 5, ncol = 6)
  sample_sizes = 50

@@

```{r, cache = TRUE, eval = FALSE}
set.seed(495)
gamma_data1 <- collect_data_gamma(10000)
write.csv(gamma_data1,"gamma_data1_10000.csv", row.names = FALSE)

normal_data1 <- collect_data_normal(10000)
write.csv(normal_data1,"normal_data1_10000.csv", row.names = FALSE)

log_normal_data1 <- collect_data_log_normal(10000)
write.csv(log_normal_data1,"log_normal1_data_10000.csv", row.names = FALSE)

beta_data1 <- collect_data_beta(10000)
write.csv(beta_data1,"beta_data1_10000.csv", row.names = FALSE)
```

```{r, eval = FALSE}
    gamma_data1 <- gamma_data5 %>%
      pivot_longer(cols = 2:6, names_to = "method", values_to = "prop")

    gamma_fig <- ggplot(data = gamma_data1, aes(x = Sample_Size, y = prop, 
                                                color = method))+
      geom_point() +
      geom_line() + 
      labs(title = "Proportion of CIs that Contain the True Parameter 
           from Gamma Distibution", x = "Sample Size", 
           y = "Proportion (of xyz Simulations)", color = "Method") +
      theme_light()

    gamma_table <- gamma_data %>%
      kable(booktabs = TRUE, align = "c", caption = "gamma", 
            col.names = gsub("[_]", " ", names(gamma_data)))

    gamma_table
    gamma_fig

@@

>>>>

---
title: A Comparision of Bootstrap Confidence Interval Methods

@@

keywords:
- simulation
- percentile
- bias-corrected
- acceleration
- studentized

@@

```{r, cache = TRUE}
data(SnowGR)
favstats(~ Dec, data = SnowGR)
gf_histogram(~ Dec, data = SnowGR) +
  labs(x = "Annual Snowfall in December", 
       title = "Distribution of Annual Snowfall in December", y = "Count",
       caption = "Fig. 1: Histogram with Studentized Interval Shown")
```

@@

```{r, cache = TRUE}
set.seed(495)
orig_mean <- mean(~ Dec, data = SnowGR) # theta hat

@@

```{r, cache = TRUE}
gf_dhistogram(~ mean, data = dec_means) %>%
  gf_dens() %>%
  gf_labs(title = "Histogram of 10,000 Bootstrapped Mean Dec Values",
          caption = "Fig. 2: Bootstrap Histogram")
```

@@

```{r, cache = TRUE}
mean(~ Dec, data = SnowGR) + qnorm(c(0.025, 0.975)) * 
  sd(~ Dec, data = SnowGR)/sqrt(nrow(SnowGR))
```

```{r, cache = TRUE}
gf_histogram(~ Dec, data = SnowGR) +
  labs(x = "Annual Snowfall in December", 
       title = "Distribution of Annual Snowfall in December", y = "Count",
       caption = "Fig. 3: Histogram with Standard Interval Shown") +
  geom_vline(aes(xintercept= 13.85639)) +
  geom_vline(aes(xintercept= 17.65957))
```

@@

```{r, cache = TRUE}
qdata(~ mean, c(0.025, 0.975), data = dec_means)
```

```{r, cache = TRUE}
gf_histogram(~ Dec, data = SnowGR) +
  labs(x = "Annual Snowfall in December", 
       title = "Distribution of Annual Snowfall in December", y = "Count",
       caption = "Fig. 4: Histogram with Percentile Interval Shown") +
  geom_vline(aes(xintercept= 13.91170)) +
  geom_vline(aes(xintercept= 17.72353))
```
This interval is similar to the one created using the Standard method, but shifted slightly to the right.

@@

```{r, cache = TRUE}
sample_mean <- mean(~Dec, data = SnowGR); sample_mean
```

Then we find the proportion of bootstrap replications that have a sample mean less than the original sample mean.
```{r, cache = TRUE}
less_than_sample_mean <- sum(ifelse(dec_means <= sample_mean, 1, 0))/10000
less_than_sample_mean
```

From this, we can calculate the bias-correction value, $z_0$:
```{r, cache = TRUE}
z0 <- qnorm(less_than_sample_mean); z0
```

@@

```{r, cache = TRUE}
alphalow <- 0.025; alphahigh <- 0.975

@@

```{r, cache = TRUE}
gf_histogram(~ Dec, data = SnowGR) +
  labs(x = "Annual Snowfall in December", 
       title = "Distribution of Annual Snowfall in December", y = "Count",
       caption = "Fig. 5: Histogram with BC Interval Shown") +
  geom_vline(aes(xintercept= 14.01172)) +
  geom_vline(aes(xintercept= 17.82017))

@@

```{r, cache = TRUE}
theta <- function(x){mean(x)}
jackknife_results <- bootstrap::jackknife(SnowGR$Dec, theta) 
```

@@

```{r, cache = TRUE}
jackmean <- mean(~ jackknife_results$jack.values); jackmean
```

@@

```{r, cache = TRUE}
estimated_a <- (1/6)*sum((jackknife_results$jack.values - jackmean)^3)/
  (sum((jackknife_results$jack.values - jackmean)^2))^(1.5); estimated_a
```

@@

```{r, cache = TRUE}
a_newlow <- pnorm(z0 + (z0 + qnorm(alphalow))/
                    (1-estimated_a*(z0 + qnorm(alphalow)))); #a_newlow
a_newhigh <- pnorm(z0 + (z0 + qnorm(alphahigh))/
                     (1-estimated_a*(z0 + qnorm(alphahigh)))); #a_newhigh
qdata(~ mean, c(a_newlow, a_newhigh), data = dec_means)
```

```{r, cache = TRUE}
gf_histogram(~ Dec, data = SnowGR) +
  labs(x = "Annual Snowfall in December", 
       title = "Distribution of Annual Snowfall in December", y = "Count",
       caption = "Fig. 6: Histogram with BCa Interval Shown") +
  geom_vline(aes(xintercept= 13.91591)) +
  geom_vline(aes(xintercept= 17.72353))

@@

```{r, cache = TRUE}
set.seed(495)

@@

This would give us the following histogram:

```{r, cache = TRUE}
gf_histogram(~ Dec, data = SnowGR) +
  labs(x = "Annual Snowfall in December", 
       title = "Distribution of Annual Snowfall in December", y = "Count", 
       caption = "Fig. 7: Histogram with Studentized Interval Shown") +
  geom_vline(aes(xintercept= 14.03467)) +
  geom_vline(aes(xintercept= 18.01709))

@@

```{r run simulation log normal, include = FALSE}

#Third, we sample from a Log-Normal distribution with mean log of 0 and a mean standard deviation of 1. Therefore, the true population mean of this distribution is $e^{\frac{1}{2}}$ [@Guillaume21].

run_simulation_log_normal <- function(simulation_rep){
  trueparameter <- exp(1/2)

@@

```{r run simulation beta, include = FALSE}
run_simulation_beta <- function(simulation_rep){
  trueparameter <- 0.8

@@

```{r collect data gamma}
collect_data_gamma <- function(sim_reps) {
  data2 <- matrix(nrow = 5, ncol = 6)
  sample_sizes = 10

  for (i in 1:5) {
    simulation_rep <- seq(1,sim_reps)
    results <- mclapply(simulation_rep, run_simulation_gamma, 
                        mc.cores = detectCores())
    df <- data.frame(matrix(unlist(results), nrow=length(results), byrow=TRUE))
    ci_prop <- colSums(df)/sim_reps

    data2[i, 1] <- sample_sizes
    data2[i, 2:6] <- ci_prop
    sample_sizes <- sample_sizes+10
  }

@@

```{r collect data log normal, include = FALSE}
collect_data_log_normal <- function(sim_reps) {
  data2 <- matrix(nrow = 5, ncol = 6)
  sample_sizes = 10

@@

    data2[i, 1] <- sample_sizes
    data2[i, 2:6] <- ci_prop
    sample_sizes <- sample_sizes+10
  }

@@

```{r collect data normal}
collect_data_normal <- function(sim_reps) {
  data2 <- matrix(nrow = 5, ncol = 6)
  sample_sizes = 10

@@

```{r collect data beta, include = FALSE}
collect_data_beta <- function(sim_reps) {
  data2 <- matrix(nrow = 5, ncol = 6)
  sample_sizes = 10

@@

```{r, cache = TRUE, eval = FALSE}
set.seed(495)
sim_reps <- 10000

gamma_data <- collect_data_gamma(sim_reps)
write.csv(gamma_data1,"gamma_data1.csv", row.names = FALSE)

normal_data <- collect_data_normal(sim_reps)
write.csv(normal_data1,"normal_data1.csv", row.names = FALSE)
```

```{r, include = FALSE, eval = FALSE}
#beta_data11 <- collect_data_beta(sim_reps)
#write.csv(beta_data1,"beta_data1_100002.csv", row.names = FALSE)

log_normal_data <- collect_data_log_normal(sim_reps)
write.csv(log_normal_data1,"log_normal_data1.csv", row.names = FALSE)

lognormal_df_long <- lognormal_df %>%
  pivot_longer(cols = 2:6, names_to = "method", values_to = "prop")

lognormal_fig <- ggplot(data = lognormal_df_long, 
                    aes(x = Sample_Size, y = prop, color = method)) + 
  geom_point() + geom_line() + 
  labs(title = "Proportion of CIs that Contain the True Parameter from Log-Normal 
       Distibution", x = "Sample Size", 
       y = "Proportion (of 10,000 Simulations)", color = "Method") +
  theme_light() + 
  theme(legend.position="bottom")

lognormal_table <- lognormal_df %>%
  kable(booktabs = TRUE, align = "c", caption = "log normal", 
        col.names = gsub("[_]", " ", names(lognormal_df)))

lognormal_df <- read.csv(file = 'log_normal1_data_10000.csv')

@@

Then, we can plot the results and compile the results into tables.

```{r create graphs}

gamma_df <- read.csv(file = 'gamma_data1_10000.csv')
normal_df <- read.csv(file = 'normal_data1_10000.csv')

gamma_df_long <- gamma_df %>%
  pivot_longer(cols = 2:6, names_to = "method", values_to = "prop")

gamma_fig <- ggplot(data = gamma_df_long, 
                    aes(x = Sample_Size, y = prop, color = method)) + 
  geom_point() + geom_line() + 
  labs(title = "Proportion of CIs that Contain the True Parameter from Gamma 
       Distibution", x = "Sample Size", 
       y = "Proportion (of 10,000 Simulations)", color = "Method",
       caption = "Fig. 8: Graph showing the Accuracy of Bootstrap CI Methods for different sample sizes on Gamma Dist.") +
  theme_light() +
  theme(legend.position="bottom")

gamma_table <- gamma_df %>%
  kable(booktabs = TRUE, align = "c", caption = "Sampled From Gamma Dist.", 
        col.names = gsub("[_]", " ", names(gamma_df)))

normal_df_long <- normal_df %>%
  pivot_longer(cols = 2:6, names_to = "method", values_to = "prop")

normal_fig <- ggplot(data = normal_df_long, 
                    aes(x = Sample_Size, y = prop, color = method)) + 
  geom_point() + geom_line() + 
  labs(title = "Proportion of CIs that Contain the True Parameter from Normal 
       Distibution", x = "Sample Size", 
       y = "Proportion (of 10,000 Simulations)", color = "Method",
       caption = "Fig. 8: Graph showing the Accuracy of Bootstrap CI Methods for different sample sizes on Normal Dist.") +
  theme_light() +
  theme(legend.position="bottom")

normal_table <- normal_df %>%
  kable(booktabs = TRUE, align = "c", caption = "Sampled from Normal Dist.", 
        col.names = gsub("[_]", " ", names(normal_df)))
```

```{r}
gamma_fig
gamma_table
```

When we sample from the Gamma distribution which is a distribution that is heavily skewed, the Studentized Confidence interval performs the best. At a large sample size, the Studentized Confidence interval contains the true population parameter about 94.5% of the time. The Studentized confidence interval outperforms the rest of the intervals by a bit. The rest of the intervals are very similar to each other ranging from a correct interval 92.5% to 93% of the time. In general, as the sample size that we are bootstrapping from increases, the higher percentage our confidence intervals predict correctly.

```{r}
normal_fig
normal_table
```

Similar to when we sampled from the Gamma distribution, the Studentized confidence interval performed the best when we sampled from a Normal distribution with around 95% of the confidence intervals containing the true population parameter. The next best is the Standard confidence interval with a performance of around 94.5% correct intervals at a high sample size. The rest of the intervals performed similarly at around 94.2% accuracy. As before, the accuracy increased as the sample size increased.
